---
title: "R Notebook"
output: html_notebook
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(sensobol)
library(tidyverse)
library(ggplot2)
library(plotly)
library(gridExtra)

set.seed(100)
```

In this notebook, the `sensobol` package is analyzed in detail. 
The computation of first-order and total-order variance-based sensitivity indices requires two elements:

1. a sampling design, that means how the sample is distributed in the multidimensional space of the model parameter,

2. estimators, formulas to compute the sensitivity indices.

All the estimators are sample-based.


```{r}
params <- c("x", "y")
N <- 2^8
a <- plot(sobol_matrices(N = N, params = params, type = "QRN"), 
     pch = 20)
```

```{r}
params <- c("x", "y")
N <- 2^8
b <- plot(sobol_matrices(N = N, params = params, type = "R"), 
     pch = 20)
```
```{r}
params <- c("x", "y")
N <- 1000

# Create a function to generate the plot
create_sobol_plot <- function(type) {
  data <- as.data.frame(sobol_matrices(N = N, params = params, type = type))
  ggplot(data, aes(x = x, y = y)) +
    geom_point(size = 1) +
    theme_minimal() +
    ggtitle(paste(type)) +
    coord_fixed(ratio = 1)  # This ensures square aspect ratio
}

# Create the plots
a <- create_sobol_plot("QRN")
c <- create_sobol_plot("LHS")

# Arrange the plots side by side
grid.arrange(a, c, ncol = 2)
```

